# Aggregate and Merge the data {#aggregate_merge}

## Aggregate the UBER Data

Remember the column `maxspeed` from the `OSM_sf` table ? We are going to combine the `DT.Uber` table with the `OSM_sf` to compare the Uber `speed_kph_mean` with the OSM `maxspeed``:

first we select only from the `OSM_sf` the fields of interest :
```{r}
DT.max.speed <- OSM_sf %>% dplyr::select(osm_id,fclass,name,maxspeed)
```

then we set this new table as a `data.table`, convert the column `osm_id` as numeric (because the `osm_way_id` from `DT.Uber` is already a numeric value, and finally define the index column :
```{r}
setDT(DT.max.speed)
DT.max.speed[,osm_id:=as.numeric(osm_id)]
setkey(DT.max.speed,osm_id)
```

And now, we merge this `DT.max.speed` table with the `DT.Uber` :
```{r}
DT.max.speed <- merge(DT.max.speed,DT.Uber, by.x='osm_id',by.y='osm_way_id')
```

with this new table we add a new column `is.over` that gives 1 if the _mean_speed_kph_ is greater than the _maxspeed_ allowed and 0 otherwise :
```{r}
DT.max.speed[,is.over := as.numeric(is.over <- speed_kph_mean>maxspeed)]
```

Out of sheer curiosity, how often is the speed limit respected ?
```{r}
DT.max.speed[,(ratio=sum(is.over)/.N)*100, ]
```
OK.


Now, we check how many ways have a speed limit defined :
```{r}
DT.max.speed[,(ratio=sum(maxspeed>0)/.N)*100, ]
```
How does it looks like per maxspeed value :
```{r}
DT.max.speed[, .(count = .N), by = .(maxspeed)][, percentage := count / sum(count) *
                                                  100][order(maxspeed)]
```
Fine, there are few ways without speed limit. (less than 0.09%).


Hang on, we are close the end, we still want to aggregate the data per way and hour of the day. For this, we are going to use the *group by* functionality via the `by` keyword, and create new variables :
(take note that we filter the table on the ways that have an OSM maximum speed defined)
```{r}
DT.max.speed.per.way.hour <-
  DT.max.speed[maxspeed > 0][, .(
    total_rows = .N,
    ratio.over = sum(is.over) / (.N) * 100,
    speed_kph_mean = speed_kph_mean <-
      mean(speed_kph_mean),
    maxspeed = maxspeed <-
      mean(maxspeed),
    speed_minus_max = speed_kph_mean - maxspeed,
    name = first(name)
  ), by = .(osm_id, hour)]
```

How does this new aggregated tables looks like ? (The empty square brackets are a neat `data.table` trick to print it)
```{r}
DT.max.speed.per.way.hour[]
```
For each OSM way, we have all hours of the day, the percentage of times the median speed was over the maximum speed allowed, and the delta between both.


