# Visualizations with ggplot2 {#viz_1}

We can plot a sample of this table to have an overview by filtering on the _hour_ equal to 17h :
```{r}
ggplot(DT.max.speed.per.way.hour[, name := as.factor(name)][hour == 17], aes(ratio.over, total_rows)) + geom_point(shape =                                                                                                                     21, alpha = .4, aes(color = (name))) + facet_wrap(. ~ maxspeed) + theme(legend.position = 'None')
```

Here we just confirm what we saw in Chapter \@ref(aggregate), it seems some _maxspeed_ are not "standard" or very seldomly used by Uber.
We can confirm this with a quick histogram :
```{r}
ggplot(DT.max.speed.per.way.hour,aes(maxspeed))+ geom_histogram()+theme_bw()
```
Indeed, 30,50 and 80 kph limits are our numbers.  
We can drop the others rows :
```{r}
DT.max.speed.per.way.hour <- DT.max.speed.per.way.hour[maxspeed %in% c(30,50,80)]
```

Now focus on the maxspeed=50 kph:(the color represents a street name).
Please notice here that we use the `ggplotly` function from the `plotly` package to make our `ggplot2` chart interactive (under the hood, it is _d3.js_ running, you can be proud of you, really.)
```{r}
p <-
  ggplot(DT.max.speed.per.way.hour[, name := as.factor(name)][hour == 17 &
                                                                maxspeed == 50 &
                                                                ratio.over < 15 &
                                                                ratio.over > 0], aes(ratio.over, total_rows)) + geom_point(shape = 1, alpha =
                                                                                                                             1, aes(color = (name))) + theme(legend.position = 'None')
ggplotly(hide_legend(p))
```

What those curves could be ? How can those dots, which color corresponds to a street name, form inverse curves  although they are not related ?  For instance, the _Müllerstraße_ dot on the plot is next to the _Frankfürter Allee_ dot, and on the same fitting curve, although they are in two different districts of Berlin.


More prosaically, we can plot the over speed percentage in the day :
For this, we just use our current table, and aggregate per hour of the day and compute the weighted ratio of over speed :
```{r}
DT.max.speed.per.hour <- DT.max.speed.per.way.hour[,.(ratio.over=sum(ratio.over*total_rows)/sum(total_rows),mean_speed_minus_max=sum(speed_minus_max*total_rows)/sum(total_rows)),by=.(hour,maxspeed)][,maxspeed:=as.factor(maxspeed)]

DT.max.speed.per.hour[0:5]
```

We use here the `ggthemes` package to add some elegance and credibility to our charts with the help of _The Economist_ theme :
```{r}
ggplot(DT.max.speed.per.hour, aes(hour, ratio.over)) + geom_line(aes(color =
                                                                       maxspeed), size = 1) + theme_economist() + ggtitle('speed infrigment % per hour of the day') +
  scale_color_economist() + theme(plot.title = element_text(size = 12)) +
  ylab('%') + xlab('hour of the day')

ggplot(DT.max.speed.per.hour, aes(hour, mean_speed_minus_max)) + geom_line(aes(color =
                                                                                 maxspeed), size = 1) + theme_economist() + ggtitle('difference between mean speed and max allowed speed per hour of the day') +
  scale_color_economist() + theme(plot.title = element_text(size = 12)) +
  ylab('difference in kph') + xlab('hour of the day')
```

Hey, but we forgot to look at the overall rides along the hour of the day !
How could we do that ? Just reuse the original `DT.Uber` table and do some aggregation :
```{r}
DT.rides.per.hour <- DT.Uber[,.(total_rides=.N),by=.(hour)]
DT.rides.per.hour[0:5]
```

For the plot, we use an other `ggplot` trick ; the **polar coordinates** :
```{r}
ggplot(DT.rides.per.hour, aes(hour, total_rides)) + geom_line(color = 'darkblue') +
  theme_economist() + ggtitle('Sum of rides per hour along the day', subtitle = 'Warning : based on already aggregated data from Uber') +
  scale_color_economist() + theme(
    plot.title = element_text(size = 12),
    legend.position = 'None',
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.ticks.x = element_blank()
  ) + ylab('') + xlab('') + coord_polar() 
```

Which gives us a relative idea of the rides volume in time.