
# Visualizations with Leaflet I{#viz_3}

## Plot densitiy of rides

Let's have a look at the distribution per hour of the day of the number of aggregated trips :

```{r}
ggplot(OSM_sf.with.Uber,aes( as.factor(hour),total_rows)) + geom_boxplot()+ theme_economist()
```
So many outliers...


The [`leaflet`](https://rstudio.github.io/leaflet/) package is the way to go when you want to implement interactive maps in R, especially when working in a [Shiny](https://shiny.rstudio.com/tutorial/) project.

we start by defining the data we want to plot: the density of trips, at 5 am for street with more than 10 occurences, not _motorway_, neither _tunnel_  :
```{r}
data.to.plot <-
  OSM_sf.with.Uber %>% filter(hour == 2   &
                                fclass != 'motorway' & tunnel == 'F')

colorPalette <-
  colorQuantile("plasma",
                data.to.plot$total_rows,
                n = 5,
                reverse = T)
```


And here we go with Leaflet : (we filter on 5 am and use a CartoDB theme, more info [here](https://leaflet-extras.github.io/leaflet-providers/preview/))


Now let's add the coordinates of the nightclubs in Berlin. For this, we load an other shapefile, the one with the buildings informations, from the GeoFabrik folder :

```{r}
OSM_sf_building  <- sf::read_sf(dsn = "data/berlin_osm_shp/", layer = "gis_osm_pois_free_1")
```

Let's filter the  night life amenities :
```{r}
OSM_sf_building.night <- OSM_sf_building %>% filter(fclass %in% c("bar","bbq","biergarten","cafe","drinking_water","fast_food","food_court","pub","restaurant","brothel","casino","cinema","nightclub","stripclub","swingerclub","biergarten"))

ggplot(OSM_sf_building.night,aes(fclass))+ geom_bar(fill='steelblue')+theme_economist()+coord_flip()
```

We add the _nightclub_ class data to the Leaflet plot :
```{r}
leaflet(OSM_sf_building.night %>% filter(fclass %in% c('nightclub')))    %>% addProviderTiles(providers$CartoDB.DarkMatter) %>% addCircleMarkers(
  stroke = F,
  fill = T,
  fillOpacity = 0.3,
  weight = 0.5,
  radius = 3,
  color = 'white',
  popup =  ~ name,
  label =  ~ name,
  layerId = OSM_sf_building.night$osm_id) %>%  addPolylines(data=data.to.plot,
                                        color = ~colorPalette(data.to.plot$total_rows),
                                        weight = 1,
                                        layerId = OSM_sf.with.Uber$osm_id)%>% addLegend(data=data.to.plot,"bottomright", pal = colorPalette, values = ~total_rows,
    title = "Rides density quantiles",
    labFormat = labelFormat(prefix = ""),
    opacity = 1
  )
```